Covariate analysis of UFL-Mayo-ISB cerebellum arrays

### 2016-01-25 code/Rmd/Mayo_Cerebellum_Array.Rmd
Added markdown for processing cerebellum microarray data from mayo-ufl-isb

---
title: "Covariate analysis of UFL-Mayo-ISB temporal cortex arrays"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./Mayo_TemporalCortex_Array.Rmd",
                                 parentId ="syn5570325",
                                 entityName = 'UFL_Mayo_ISB Temporal Cortex Microarray')
```

@@

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = FALSE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570325';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of UFL_Mayo_ISB temporal cortex array data';

thisFileName <- 'Mayo_TemporalCortex_Array.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
EXPR_ID <- 'syn3617054';
ALL_USED_IDs <- EXPR_ID
EXPR_OBJ <- synGet(EXPR_ID)
EXPR <- fread(EXPR_OBJ@filePath,sep=',',data.table=F)

# Get clinical metadata
METADATA_ID <- 'syn3617056'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_ID
METADATA_OBJ <- synGet(METADATA_ID)
METADATA <- read.table(METADATA_OBJ@filePath,sep=',',header=T, stringsAsFactors=F)
```

@@

METADATA = dplyr::mutate(METADATA,
                         plate2 = 2*plate2,
                         plate3 = 3*plate3,
                         plate4 = 4*plate4,
                         plate5 = 5*plate5,
                         plate = plate2+plate3+plate4+plate5)
METADATA$plate[METADATA$plate == 0] = 1
```
Temporal cortex of `r dim(EXPR)[2]` subjects from the Mayo cohorts were obtained from [AMP-AD portal](`r EXPR_ID`)

@@

tree = hclust(as.dist(t(EXPR)))
plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
                    dendroLabels = FALSE,
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
```
### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, results='asis', echo=FALSE}
# Find correlation between PC's of gene expression with covariates
DM = getDesignMatrix(COVARIATES, Intercept=F)
DESIGN = getDesignMat2Fact(DM$design,FactorCovariates)
colnames(DESIGN) = substr(colnames(DESIGN),1,50)
preAdjustedSigCovars = runPCAandPlotCorrelations(EXPR, DESIGN,
                                                 'NULL design',
                                                 isKeyPlot=TRUE)

@@

# Query biomart
Illumina2HGNC <- getBM(attributes = c("illumina_humanht_12_v4", "hgnc_symbol"),
                       filters = "illumina_humanht_12_v4", 
                       values = RESIDUAL.GENE_EXPRESSION$illumina_humanht_12_v4,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(Illumina2HGNC, RESIDUAL.GENE_EXPRESSION) %>%
  dplyr::select(illumina_humanht_12_v4, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Temporal Cortex Microarray', parentId = parentId)
CODE <- synStore(CODE)

>>>>

---
title: "Covariate analysis of UFL-Mayo-ISB cerebellum arrays"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(knit2synapse)
library(synapseClient)

@@

knit2synapse::knitToFolderEntity(file = "./Mayo_Cerebellum_Array.Rmd",
                                 parentId ="syn5570325",
                                 entityName = 'UFL_Mayo_ISB Cerebellum Microarray')
```

@@

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE)
```

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn5570325';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of UFL_Mayo_ISB Cerebellum array data';

thisFileName <- 'Mayo_Cerebellum_Array.Rmd'

# Github link
thisRepo <- getRepo(repository = "th1vairam/Brain_Reg_Net", ref="branch", refName='AMPAD')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('code/Rmd/',thisFileName))
```
### Data download
#### Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
EXPR_ID <- 'syn3256501';
ALL_USED_IDs <- EXPR_ID
EXPR_OBJ <- synGet(EXPR_ID)
EXPR <- fread(EXPR_OBJ@filePath,sep=',',data.table=F)

# Get clinical metadata
METADATA_ID <- 'syn3256502'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = METADATA_ID
METADATA_OBJ <- synGet(METADATA_ID)
METADATA <- read.table(METADATA_OBJ@filePath,sep=',',header=T, stringsAsFactors=F)
```

@@

METADATA = dplyr::mutate(METADATA,
                         plate0 = 6*plate0,
                         plate1 = 1*plate1,
                         plate2 = 2*plate2,
                         plate3 = 3*plate3,
                         plate4 = 4*plate4,
                         plate = plate0+plate1+plate2+plate3+plate4)
METADATA$plate[METADATA$plate == 0] = 5
METADATA$plate[METADATA$plate == 6] = 0
```
Cerebellum of `r dim(EXPR)[2]` subjects from the Mayo cohorts were obtained from [AMP-AD portal](`r EXPR_ID`)

@@

tree = hclust(as.dist(t(EXPR)))
plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
#                     dendroLabels = FALSE,
                    abHeight = 10, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
```
### Significant Covariates
Correlation between pca of unadjusted mRNA expression and covariates is used to find significant covariates
```{r preAdjusted.covariates, results='asis', echo=FALSE}
# Find correlation between PC's of gene expression with covariates
DM = getDesignMatrix(COVARIATES, Intercept=F)
DESIGN = getDesignMat2Fact(DM$design,FactorCovariates)
colnames(DESIGN) = substr(colnames(DESIGN),1,50)
preAdjustedSigCovars = runPCAandPlotCorrelations(EXPR, DESIGN,
                                                 'NULL design',
                                                 isKeyPlot=TRUE)

@@

# Query biomart
Illumina2HGNC <- getBM(attributes = c("illumina_humanht_12_v4", "hgnc_symbol"),
                       filters = "illumina_humanht_12_v4", 
                       values = RESIDUAL.GENE_EXPRESSION$illumina_humanht_12_v4,
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(Illumina2HGNC, RESIDUAL.GENE_EXPRESSION) %>%
  dplyr::select(illumina_humanht_12_v4, hgnc_symbol, one_of(rownames(COVARIATES)))
```
### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'UFL_Mayo_ISB Cerebellum Microarray', parentId = parentId)
CODE <- synStore(CODE)
